---
title: "Covid91 vaccine study Final2023"
author: "Kevin Brandon Salazar"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```


# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

$H_0$ The vaccine is not effective in males for preventing covid

$H_A$ The vaccine is effective in males for preventing covid

# Males

## Methods

The response method is Cat~Cat, because the question is about the effectiveness of the vaccine in males therefore we have to subset the data to include males only. 

Graphical display bar chart, Numerical display xtabs, Inferential display $x^2$ test

## Results

```{r}
Males = filter(FinalData, sex == "M")
```

## Descriptive Results
## Graphical Descriptive Results

```{r}
barchartGC(~infected+treatment,data=Males)
barchartGC(~infected+treatment,data=Males, type="percent")

```

This graph shows among the people who did get the vaccine while they were infected (covid19) showed a difference. Meanwhile among those who were not infected (OK) and took the vaccine no difference was showed.

## Numerical Descriptive Results

```{r}
table1 <- xtabs(~infected+treatment, data=Males)
rowPerc(table1)
colPerc(table1)
```

Among all the people infected it appears that the people who had the placebo (70%) are 2 times more likely to have covid than the people with the vaccine (30%) this shows that the vaccine is very effective. Meanwhile among all the people not infected (OK) it appears that the people who had the placebo (50%) are barely more likely to have covid than the people with the vaccine (49.5%) this shows that the vaccine is very ineffective in this scenario.

## Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

In the Pearson's Chi-squared test P-Value is 0.0000000000001098 shows that the vaccine is effective in preventing covid among males. If you get the vaccine you are 2.28 less likely to get covid than someone who gets the placebo.

In the Fisher's Exact Test P-Value is 0000000.23 which shows we can reject the Null Hypothesis. The odds ratio is that you are 0.35 times likely to get aids when you have the placebo than the vaccine. The answer is still the same.

# Females

$H_0$ The vaccine is not effective in females for preventing covid

$H_A$ The vaccine is effective in females for preventing covid

## Methods

The response method is Cat~Cat, because the question is about the effectiveness of the vaccine in females therefore we have to subset the data to include females only. 

Graphical display bar chart, Numerical display xtabs, Inferential display $x^2$ test

## Results

```{r}
Females = filter(FinalData, sex == "F")
```

## Descriptive Results
## Graphical Descriptive Results

```{r}
barchartGC(~infected+treatment,data=Females)
barchartGC(~infected+treatment,data=Females, type="percent")

```

## Numerical Descriptive Results

```{r}
table2 <- xtabs(~infected+treatment, data=Females)
rowPerc(table2)
colPerc(table2)
```

## Inferential Results

```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)
```

# LGBTQ

$H_0$ The vaccine is not effective in LGBTQ for preventing covid

$H_A$ The vaccine is effective in Elicit Drug Users for preventing covid


## Methods

The response method is Cat~Cat, because the question is about the effectiveness of the vaccine in LGBTQ therefore we have to subset the data to include LGBTQ only. 

Graphical display bar chart, Numerical display xtabs, Inferential display $x^2$ test

## Results

```{r}
LGBTQ = filter(FinalData, LGBTQ == "gay")
```

## Descriptive Results
## Graphical Descriptive Results

```{r}
barchartGC(~infected+treatment,data=LGBTQ)
barchartGC(~infected+treatment,data=LGBTQ, type="percent")

```

## Numerical Descriptive Results

```{r}
table3 <- xtabs(~infected+treatment, data=LGBTQ)
rowPerc(table3)
colPerc(table3)
```

## Inferential Results

```{r}
chisq.test(table3)
chisqtestGC(table3)
fisher.test(table3)
```

# Druggies

$H_0$ The vaccine is not effective in Elicit Drug Users for preventing covid

$H_A$ The vaccine is effective in Elicit Drug Users for preventing covid


## Methods

The response method is Cat~Cat, because the question is about the effectiveness of the vaccine in Druggies therefore we have to subset the data to include Druggies only. 

Graphical display bar chart, Numerical display xtabs, Inferential display $x^2$ test

## Results

```{r}
DrugUser = filter(FinalData, DrugUser == "yes")
```

## Descriptive Results
## Graphical Descriptive Results

```{r}
barchartGC(~infected+treatment,data=DrugUser)
barchartGC(~infected+treatment,data=DrugUser, type="percent")

```

## Numerical Descriptive Results

```{r}
table4 <- xtabs(~infected+treatment, data=DrugUser)
rowPerc(table4)
colPerc(table4)
```

## Inferential Results

```{r}
chisq.test(table4)
chisqtestGC(table4)
fisher.test(table4)
```

# Overall Results and Conclusions